API design for machine learning software: experiences from the scikit-learn project

### 2013-06-24 paper.tex
rework the conclusion and separate out future work

We have discussed the scikit-learn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We motivated choices and compromises made in its design.
We argue that the API is powerful, usable and extensible,
as witnessed by the appearance of third-party packages
that follow the scikit-learn conventions and tie into the toolkit.\footnote{
  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
  this mention does not imply endorsement of this package.}
While part of the scikit-learn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.

A \sklearn~1.0 release is nearing,
but several problems must still be resolved.
At present, the library does not support some classical machine learning tasks,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Topics like structured prediction or reinforcement learning are
considered off-topic for the project, since they would require quite different
data representations and APIs.

@@

>>>>

\section{Future Work}
A \sklearn~1.0 release is nearing,
but several problems must still be resolved.
At present, the library does not support some classical machine learning tasks,
including neural networks, ensemble meta-estimators for
bagging or subsampling strategies and missing value completion algorithms.
Topics like structured prediction or reinforcement learning are
considered off-topic for the project, since they would require quite different
data representations and APIs.

@@

We have discussed the scikit-learn API
and the way it maps machine learning concepts and tasks
onto objects and operations in the Python programming language.
We have shown how a consistent API across the package makes scikit-learn
very \textbf{usable} in practice: experimenting with different learning
algorithm is as simple as substituting a new class definition.
Through composition interfaces such as Pipelines, Feature Unions,
and meta-estimators, these simple building blocks lead to an API which is
\textbf{powerful}, and can accomplish a wide variety of learning tasks
within a small amount of easy-to-read code.
Through duck-typing, the consistent API leads to a framework that is
easily \textbf{extensible}, and allows user-defined estimators to be
incorporated into the scikit-learn workflow without any explicit object
inheritance.

The simplicity, power, and extensibility of the scikit-learn API is evidenced
by the large and growing user-base, the use of scikit-learn to solve real
problems across a wide array of fields,
as well as the appearance of third-party packages
that follow the scikit-learn conventions and tie into the toolkit.\footnote{
  E.g.\ nolearn, \url{http://pythonhosted.org/nolearn};
  this mention does not imply endorsement of this package.}
While part of the scikit-learn API is necessarily Python-specific,
core concepts may be applicable to
machine learning applications and toolkits
written in other (dynamic) programming languages.
